Handling String-Number Conversion in String Constraint Solving

### 2019-11-14 draft.tex
fix a compile bug
property of product, just for discussion
another version of Lemma 5.4 for PA x PA
typo
note below another version of L 5.4

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
Encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints. 
%
%\newcommand\restrict[1]{\raisebox{-.5ex}{$\big|$}_{#1}}
\begin{lemma}\label{lemma:synccorrect}
%$\semof {\exists V_\times . \syncfof A {A'}}\restrict{\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
$\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}
%
Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the modes of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $V_\times = (\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we have to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

@@

\todo{Add an example: $\sti{x}=\sti{y} \wedge x= y\cdot y$. Do product construction informally here.}

>>>>

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
Encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints. 
%
\begin{lemma}\label{lemma:synccorrect}
%$\semof {\exists V_\times . \syncfof A {A'}}\restrict{\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
$\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}
%
Another version:
\begin{lemma}
$
\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = 
\{
(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
x\in L(A), x'\in L(A'), I(x) = I(x')
\}
$
\end{lemma}
Notice that if $A$ is flat (or, symmetrically, if $A'$ is flat), then $\semof A \cap \semof {A'}$ can be still decoded from $\semof {\syncfof A {A'}}$ due to Corollary~\ref{corollary:pfa}, 
namely in the form of\\ $\semof A \cap \semof {A'} = \{\decode A(I\cup x) \mid \exists J.((I\cup x) \cup J) \in \semof{\syncfof A {A'}}\}$.  
\bigskip

Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the modes of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $V_\times = (\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we have to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

@@
